---
title: "GenomeAdapt: Detecting Signatures of Local Adaptation Based on Ancestry Trajectories"

author: "Xinghu Qin"
#date: "`r Sys.Date()`"
date: "11-09-2021"
URL: https://xinghuq.github.io/GenomeAdapt/index.html
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Instruction for Package GenomeAdapt }
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Instruction

This vignette demonstrates the implementation of GenomeAdapt for detecting signatures of local adaptation based on multidimensional ancestry trajectories (_n_ * _n_ ancestry genetic trajectories, n is the number of individuals). 
If n samples are included in the analysis, there will be n dimensional spaces that represent the common ancestry maps based on the identity-by-descent (IBD). 

The package calculates the correlations of loci with the common ancestry genetic maps adopting the Genomic Data Structure (GDS, Zheng et al., 2012<doi: 10.1093/bioinformatics/bts606>) and suitable for millions of SNP data. Loci sharing a greater level of most recent common ancestor (MRCA) (large Z-scores) indicate a large number of individuals descend from recent common ancestors, which signals the rapid increase in frequency of a beneficial allele due to recent positive selection. The rationale underlying this package is somewhat analogous to KLFDAPT (Qin, 2021<DOI:10.1101/2021.05.15.444294>, https://xinghuq.github.io/KLFDAPC/articles/Genome_scan_KLFDAPC.html). It combines the concept of IBD-based genome scan (Albrechtsen et al., 2010)<doi: 10.1534/genetics.110.113977>, iHS (Voight, 2006) <doi:10.1371/journal.pbio.0040072>, and eigenanalysis of SNP data with an identity by descent interpretation (Zheng & Weir (2016)<DOI: 10.1016/j.tpb.2015.09.004>). It can also be interpreted as spatial varying selection as ancestry genetic maps reflect geographic origins.

Besides the detection of local adaptation, this package also estimates the population admixtures and plots its geographic genetic structure. 


# Practical analyses

We use the Hapmap dataset as an example to show how you can detect signatures of local adaptation. 


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

##Install GenomeAdapt

```{r, library}
#Install from CRAN
#install.packages("GenomeAdapt")
## or you can get the latest version of HierDpart from github
#library(devtools)
#install_github("xinghuq/GenomeAdapt")

library("GenomeAdapt")


```

```{r}
# example genepop file
# using Hapmap data
#Do genome scan
HapmapScan=GenomeAdapt.gds(genfile = SNPRelate::snpgdsExampleFileName(),method="EIGMIX",num.thread = 6, autosome.only=TRUE, remove.monosnp=TRUE, maf=0.01, missing.rate=0.1)

## it takes a while to finish this


```


## Estimating the ancestry proportion and plot the ancestry admixture 

We can use the IBD-based eigen analysis to estimate the ancestry proportion. Below, we show the estimation of ancestry using AdmixProt function.

```{r fig2, fig.height = 5, fig.width = 8.5, fig.align = "center"}
# get population information
library(SNPRelate)
genofile <- SNPRelate::snpgdsOpen(SNPRelate::snpgdsExampleFileName())
pop_code <- read.gdsn(index.gdsn(genofile, "sample.annot/pop.group"))

# get sample id
samp.id <- read.gdsn(index.gdsn(genofile, "sample.id"))
snpgdsClose(genofile)

# define groups
groups <- list(CEU = samp.id[pop_code == "CEU"],
    YRI = samp.id[pop_code == "YRI"],
    CHB = samp.id[is.element(pop_code, c("HCB", "JPT"))])

### estimate the ancestry proportion

Admixpro=AdmixProp(HapmapScan,groups=groups,bound=TRUE)

PlotAdmix(Admixpro,group=as.factor(pop_code), xlab="Individuals",multiplot = FALSE)

```

Fig.1. The ancestry admixture plot for Hapmap dataset.


## Plot the Manhattan plot  

We can get the z-scores for loci relating to the ancestry trajectories from GenomeAdapt object, then plot the manhattan plot to show the p-values.

```{r fig1, fig.height = 5, fig.width = 8.5, fig.align = "center"} 


## Calculating the Z-score for loci  
 Hapmapqval=zscores_qvals(HapmapScan)
## plot

 plotmanhattan(Hapmapqval$pvals$pvals$p.values,col=Hapmapqval$chr)

```

Fig.2. Manhattan plot of Hapmap dataset for detecting signals of local adaptation.



# References

Laloë, D., Jombart, T., Dufour, A.-B. & Moazami-Goudarzi, K. (2007). Consensus genetic structuring and typological value of markers using multiple coinertia analysis. Genetics Selection Evolution, 39, 545.

Qin, X., Chiang, C. W., & Gaggiotti, O. E. (2021). Kernel Local Fisher Discriminant Analysis of Principal Components (KLFDAPC) significantly improves the accuracy of predicting geographic origin of individuals. bioRxiv.

Zheng, X., & Weir, B. S. (2016). Eigenanalysis of SNP data with an identity by descent interpretation. Theoretical population biology, 107, 65-76.

Albrechtsen, A., Moltke, I., & Nielsen, R. (2010). Natural selection and the distribution of identity-by-descent in the human genome. Genetics, 186(1), 295-308.

Duforet-Frebourg, N., Luu, K., Laval, G., Bazin, E., & Blum, M. G. (2016). Detecting genomic signatures of natural selection with principal component analysis: application to the 1000 genomes data. Molecular biology and evolution, 33(4), 1082-1093.

Zheng, X., Levine, D., Shen, J., Gogarten, S. M., Laurie, C., & Weir, B. S. (2012). A high-performance computing toolset for relatedness and principal component analysis of SNP data. Bioinformatics, 28(24), 3326-3328.
